---
title: "CO2 Flux Analysis: Per Chamber, Per Date"
author: "Risper Nyairo"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Required Libraries

```{r libraries}
library(dplyr)
library(gasfluxes)
library(ggplot2)
```

## Import Data

Data come from two plots (`plotNo`, column one of the file), each with
several chambers (`chamber`), sampled on several dates (`date`), with up to
four timepoints per chamber-date (`TimeStep`: T0–T3). `Flags` marks whether a
given timepoint reading is usable (`1`) or not (`0`).

```{r import-data}
CO2data <- read.csv("AnalysisReadyCO2.csv", header = TRUE)
```

## Filter to Flag == 1

```{r filter-flags}
CO2_valid <- CO2data %>%
  filter(Flags == 1)
```

## Data Quality Check

`gasfluxes` needs a column (or set of columns) that uniquely identifies each
chamber-date series. The file already has an `ID` column for this, but it
doesn't always agree with `plotNo`/`chamber`/`date` — the check below finds
any row where the leading digit of `ID` doesn't match `plotNo`:

```{r qc-check}
CO2data %>%
  filter(substr(ID, 1, 1) != as.character(plotNo)) %>%
  select(plotNo, chamber, date, TimeStep, ID)
```

Because of that mismatch, the series identifier for flux fitting is built
directly from `plotNo`, `chamber`, and `date` instead of from `ID`, so a
mislabeled `ID` can't silently merge two different chambers' data or split
one chamber's date into two series.

## Keep Chamber-Dates with at Least 3 Timepoints

A linear flux fit needs at least 3 concentration-time points. After
filtering to `Flags == 1`, some chamber-date combinations have only 1 or 2
usable timepoints left and can't be fit — those are dropped here, and
reported so nothing disappears silently.

```{r filter-timepoints}
CO2_counts <- CO2_valid %>%
  add_count(plotNo, chamber, date, name = "n_timepoints")

dropped <- CO2_counts %>%
  filter(n_timepoints < 3) %>%
  distinct(plotNo, chamber, date, n_timepoints)

dropped

CO2_flux_ready <- CO2_counts %>%
  filter(n_timepoints >= 3)
```

## Calculate Fluxes per Chamber per Date

`.id = c("plotNo", "chamber", "date")` fits a separate flux for every
chamber on every date, rather than pooling a chamber's measurements across
dates. `.C = "CO2_conc"` is used directly since each chamber-date-timepoint
now has exactly one valid reading after the flag filter (no averaging of
duplicate readings is needed).

```{r flux-calculation, cache=TRUE}
flux.results <- gasfluxes(CO2_flux_ready,
                          .id = c("plotNo", "chamber", "date"),
                          .V = "cham_V", .A = "cham_A",
                          .C = "CO2_conc",
                          .times = "time",
                          methods = c("linear"), plot = TRUE)
```

## Results

```{r results-table}
flux.results %>%
  select(plotNo, chamber, date, linear.f0, linear.f0.se, linear.f0.p, linear.r)
```

## Daily Mean Flux per Plot

Each plot has up to 4 chambers (A–D) measured on the same date. Averaging
`linear.f0` across chambers within a plot and date collapses the per-chamber
fluxes down to a single daily mean flux per plot. `date` is also converted
from Excel's serial date format to an actual calendar date, since that's
what a time series axis should show ("Sampling Day").

```{r daily-mean-flux}
daily_flux <- flux.results %>%
  mutate(plotNo = as.factor(plotNo),
         SamplingDay = as.Date(date, origin = "1899-12-30")) %>%
  group_by(plotNo, SamplingDay) %>%
  summarise(mean_flux = mean(linear.f0, na.rm = TRUE),
            n_chambers = n(),
            .groups = "drop")

daily_flux
```

## Time Series of Daily Mean Flux

```{r flux-timeseries}
ggplot(daily_flux, aes(x = SamplingDay, y = mean_flux, color = plotNo)) +
  geom_line() +
  geom_point(size = 2) +
  theme_bw(base_size = 18) +
  theme(panel.grid = element_blank()) +
  xlab("Sampling Day") +
  ylab(expression(paste(CO[2], " flux (mg/", m^2, "/hr)"))) +
  scale_color_discrete(name = "Plot", labels = c("Fallow", "Cropped"))
```
